  
 ```{r setup, include=FALSE}
 knitr::opts_chunk$set(echo = TRUE)
 ```
 # Peer-graded Assignment: Course Project 1#
 ## Activity monitoring devices##

 ## Loading and preprocessing the data
 ### (1) get data

 ```{r echo=TRUE}
 if(!file.exists("repdata data activity.zip")) {
       download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip","repdata data activity.zip")
 }
 unzip("repdata data activity.zip")
 data <- read.csv("activity.csv")
 ```

 ### (2) Histogram of the total number of steps taken each day

 ## What is mean total number of steps taken per day?
 ```{r echo=TRUE}
 daily_sum <- aggregate(steps~date,data=data,na.rm=TRUE,sum)

 hist(daily_sum$steps,xlab="Steps/day",main="Histogram for steps per day")
 ```

 ### (3) Mean and median number of steps taken each day

 ## What is the average daily activity pattern?
 ```{r echo=TRUE}
 mean_steps <- mean(daily_sum$steps)
 median_steps <- median(daily_sum$steps)
 ```

 Mean for numbers of steps per day is `r as.character(round(mean_steps,2))` and median is `r median_steps`.

 ### (4) Time series plot of the average number of steps taken

 ## Imputing missing values
 ```{r echo=TRUE}
 interval_average <- aggregate(steps~interval,data=data,na.rm=TRUE,mean)
 wm <- which.max(interval_average$steps)
 interval_average$steps[wm]
 interval_average$interval[wm]

 library("ggplot2")
 p <- ggplot(interval_average, aes(x=interval, y=steps)) +
       geom_line() + 
       xlab("intervals")+geom_point(aes(x=interval[wm], y=steps[wm]))+geom_text(aes(x=interval[wm]+150, y=steps[wm]+10,label=paste("[",interval[wm],", ",round(steps[wm]),"]")))
 p
 ```

 ### (5) The 5-minute interval that, on average, contains the maximum number of steps

 ## Are there differences in activity patterns between weekdays and weekends?
 ```{r echo=TRUE}
 interval_average <- aggregate(steps~interval,data=data,na.rm=TRUE,mean)
 wm <- which.max(interval_average$steps)
 interval_average$steps[wm]
 interval_average$interval[wm]
 ```

 Maximal value is `r as.character(round(interval_average$steps[wm],2))` step for the interval `r interval_average$interval[wm]`.


 ### (6) Code to describe and show a strategy for imputing missing data

 I have taken into account two strategies for imputing missing values:

 1. not to change interval mean: we will replace each missing value for the mean of the corresponding interval. It means, that the mean of the interval over whole time will not change. The reason - we can expect, that the behavior of the person will be the same across the long term period.
 2. the similar to 1. - instead of mean we can use median which do not change medians of the corresponding interval over the whole period

 As global median and mean is very similar (difference = 1.19), so the approaches are almost equivalent. I will use the case 1.

 ```{r}

 library("dplyr")

 data_imp <- merge(x=data,y=interval_average, by="interval")
 NAs <- is.na(data_imp$steps.x)
 data_imp$steps.x[NAs]<-data_imp$steps.y[NAs]
 data_imp <- data_imp %>% mutate(steps=steps.x)

 ```
 In the original data there was `r sum(NAs)` rows with NA. Those were replaced by mean of the corresponding interval.


 ### (7) Histogram of the total number of steps taken each day after missing values are imputed

 ```{r echo=TRUE}
 daily_sum_imp <- aggregate(steps~date,data=data_imp,sum)

 hist(daily_sum_imp$steps,xlab="Steps/day",main="Histogram for steps per day (with imputed values)")
 ```

 ### (8) Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

 ```{r}
 library("dplyr")

 data_imp <- data_imp %>% mutate(working=(as.POSIXlt(date)$wday>0 & as.POSIXlt(date)$wday<6))
 interval_weekdays_average <- aggregate(steps~interval+working,data=data_imp,mean)

 par(mfrow=c(1,2))
 plot(type="l", x=interval_weekdays_average$interval[interval_weekdays_average$working], y=interval_weekdays_average$steps[interval_weekdays_average$working],xlab="Interval",ylab="working days",main="Time series - steps/interval")
 plot(type="l", x=interval_weekdays_average$interval[!interval_weekdays_average$working], y=interval_weekdays_average$steps[!interval_weekdays_average$working],xlab="Interval",ylab="weekend")

 ```

 ### (9) All of the R code needed to reproduce the results (numbers, plots, etc.) in the report

 DONE :)

 ### Cleaning house
 ```{r}
 rm(data,
       daily_sum,
       mean_steps,median_steps,
       interval_average, wm, p,
       data_imp, NAs,
       daily_sum_imp,
       interval_weekdays_average
    )